% =========================
% puzzles/pXX.tex
% Final - High-Card Poker: "The Claimed Jack"
% =========================

\PuzzleChapterIntro{High-Card Poker: The Claimed Jack}{The Price of a Name}{This is a Bayesian inference puzzle where the likelihood of a claim is governed by a strict institutional protocol. A specific card is named, but its reliability depends on a historical ledger modified by ``primer'' marks and a mechanical wheel that enforces false identities. The problem tests your ability to translate narrative rules into exact probabilistic weights, handle the conditioning of a lie on the available sample space, and compute the shift between a baseline and a posterior winning probability.}

\Lore{

In the Archive Casino, a named claim is treated as a financial instrument.

The high-card side duel is simple: one card each, higher rank takes the pot, ties return the stake without profit. What complicates the table is not the game, but the archive. The house maintains a ledger of named claims and will not permit a player to name a card unless the claim is governed by the published policy for that table.

A brass plate is bolted to the rail. It does not list probabilities; it lists governance.

First, the record:

\begin{quote}
\textit{Named-claim record (last 1200): truth:lie odds = 1:2.}
\end{quote}

Second, the adjustment:

\begin{quote}
\textit{Before pricing a new name, enter the Primer: one Truth and one Lie.}
\end{quote}

The table does not debate what the Primer means. The clerk amends the ledger accordingly and treats the amended fraction of Truth as the trust placed in the next named claim.

False names are regulated by mechanism rather than discretion. If a player is not speaking the card they hold, the dealer invokes the Wheel: a balanced device that selects a card-identity uniformly from all identities except the one actually held. The Wheel disregards what is already visible on the felt; it excludes only the card in hand.

}

\Problem

\textbf{High-card rules.} A standard deck has \(52\) distinct cards. Each player holds exactly one card. Ranks are ordered
\[
2<3<\cdots<10<J<Q<K<A.
\]
Suits do not matter. Higher rank wins; a tie is not a win.

\medskip

\textbf{Deal.} Raven draws \(7\clubsuit\) and leaves it face-up. Sable draws one card uniformly at random from the remaining \(51\) cards and keeps it hidden. Before any reveal, Sable announces:

\begin{quote}
``My card is \(J\diamondsuit\).''
\end{quote}

\medskip

\textbf{Named-claim policy.} Interpret the reliability of Sable's announcement, and the distribution of a false announcement, exactly as specified on the brass plate described in the Lore.

\VaultDirective{By how many percentage points does Raven's probability of a strict win change after hearing the announcement, compared to the no-announcement baseline?

Report the magnitude as a whole-number percentage in the form \(D\%~(R)\) if the announcement reduces Raven's winning chance, or \(D\%~(I)\) if it increases it.}

\SolutionPage

Let \(S\) be the event ``Sable says \(J\diamondsuit\),'' and \(A\) the event ``Sable actually holds \(J\diamondsuit\).''

\medskip

\textbf{Baseline (no announcement).}
Among the \(51\) possible hidden cards, Raven's \(7\) wins exactly against ranks \(2\) through \(6\), i.e.\ \(5\) ranks \(\times 4\) suits \(=20\) cards. Hence
\[
W_0=\frac{20}{51}\approx 0.3921569.
\]

\medskip

\textbf{Trust level from the brass plate.}
Truth:lie odds \(1:2\) over \(1200\) hands gives \(400\) truths and \(800\) lies. The Primer adds one truth and one lie, so the trust level for the next named claim is
\[
p=\frac{401}{1202}.
\]

\medskip

\textbf{Hearing \(J\diamondsuit\).}
Since Raven holds \(7\clubsuit\),
\[
P(A)=\frac{1}{51},\qquad P(\neg A)=\frac{50}{51}.
\]
Also
\[
P(S\mid A)=p,\qquad P(S\mid \neg A)=(1-p)\cdot\frac{1}{51},
\]
because in false-mode the Wheel selects uniformly among the \(51\) identities not equal to the card in hand (including identities already visible on the felt).

Bayes gives
\[
P(\neg A\mid S)=\frac{P(S\mid \neg A)P(\neg A)}{P(S\mid \neg A)P(\neg A)+P(S\mid A)P(A)}
=\frac{50(1-p)}{50(1-p)+51p}
=\frac{13350}{20167}.
\]

\medskip

\textbf{Win after the announcement.}
Raven loses if \(A\) occurs since \(J>7\). If \(\neg A\) occurs under \(S\), Sable's actual card is uniform among the \(50\) cards excluding \(J\diamondsuit\) and excluding Raven's \(7\clubsuit\). Raven wins against \(20\) of these cards, so
\[
W_1=P(\neg A\mid S)\cdot\frac{20}{50}
=\frac{13350}{20167}\cdot\frac{2}{5}
=\frac{5340}{20167}\approx 0.2647890.
\]

\medskip

\textbf{Change in percentage points.}
\[
100(W_1-W_0)\approx 26.47890-39.21569\approx -12.73679.
\]
The magnitude rounds to \(13\), and the sign is a reduction.

\FinalAnswer{$13\%~(R)$}

\NextPage