\PuzzleChapterIntro{The Divisibility \& Factorization Scroll}{Warmup \textbullet\ Möbius \textbullet\ Ramanujan \textbullet\ Cyclotomy}{This is a ``True/False'' gauntlet covering five classic identities in multiplicative number theory. You must verify properties of the Möbius function, Dirichlet convolutions, and cyclotomic polynomials to derive the correct binary string.}
\Lore{A scroll of copper foil lies on the lectern, etched with five claims.

The Archivist does not ask for proofs here.

\textit{“Only mark what is true. The Vault reads patterns, not prose.”}}

\Problem

For a positive integer $n$, let:
\begin{itemize}
  \item $\mu(n)$ be the Möbius function,
  \item $\varphi(n)$ be Euler's totient,
  \item $\sigma(n)=\sum_{d\mid n} d$ be the sum-of-divisors function,
  \item $\operatorname{rad}(n)=\prod_{p\mid n} p$ be the product of distinct primes dividing $n$.
\end{itemize}

For integers $q\ge 1$ and $n$, define the Ramanujan sum
\[
 c_q(n)=\sum_{\substack{1\le k\le q\\(k,q)=1}} e^{2\pi i kn/q}.
\]

For $n\ge 1$, define the $n$th cyclotomic polynomial
\[
\Phi_n(x)=\prod_{\substack{1\le k\le n\\(k,n)=1}}\left(x-e^{2\pi i k/n}\right).
\]

\medskip

Consider the following statements.

\begin{enumerate}[label=\textbf{(\arabic*)},leftmargin=*]
\item \textbf{(Möbius--radical transform.)} For all $n\ge 1$,
\[
\sum_{d\mid n} \mu(d)\,\operatorname{rad}(n/d)=\varphi(n)
\quad\Longleftrightarrow\quad
n\text{ is squarefree}.
\]

\item \textbf{(Möbius inversion for $\sigma$.)} For all $n\ge 1$,
\[
\sum_{d\mid n} \mu(d)\,\sigma(n/d)=n.
\]

\item \textbf{(Radical as an exponent-weighted divisor product.)} For all $n\ge 1$,
\[
\prod_{d\mid n} d^{\mu(n/d)}=\operatorname{rad}(n).
\]

\item \textbf{(Ramanujan divisor sum.)} For all $q\ge 1$ and integers $n$,
\[
\sum_{d\mid q} c_d(n)=
\begin{cases}
q,& q\mid n,\\
0,& q\nmid n.
\end{cases}
\]

\item \textbf{(Cyclotomic Möbius product.)} For all $n\ge 1$,
\[
\Phi_n(x)=\prod_{d\mid n} (x^d-1)^{\mu(n/d)}.
\]
\end{enumerate}

\medskip

\VaultDirective{Let $t_i=1$ if statement $(i)$ is true and $t_i=0$ if it is false. Output the 5-bit string $t_1t_2t_3t_4t_5$.}

\SolutionPage

\begin{enumerate}[label=\textbf{\arabic*.},leftmargin=*]

\item \textbf{True.}
Let $R=\operatorname{rad}(n)$. Since $\mu(d)=0$ unless $d$ is squarefree, only divisors $d\mid R$ contribute, and for such $d$ we have $\operatorname{rad}(n/d)=R/d$. Hence
\[
\sum_{d\mid n}\mu(d)\,\operatorname{rad}(n/d)=\sum_{d\mid R}\mu(d)\,\frac{R}{d}
=R\prod_{p\mid R}\left(1-\frac1p\right)=\varphi(R).
\]
So the left-hand side equals $\varphi(R)$ always, and it equals $\varphi(n)$ iff $\varphi(R)=\varphi(n)$.
But $\varphi(n)=\varphi(R)\prod_{p^a\| n}p^{a-1}$, so equality holds iff every $a=1$, i.e. $n$ is squarefree.

\item \textbf{True.}
This is the Dirichlet convolution identity $\mu*\sigma=\mathrm{id}$, i.e.
\[
\sum_{d\mid n}\mu(d)\,\sigma(n/d)=n.
\]

\item \textbf{False.}
Take $n=12$. Then the product is
\[
\prod_{d\mid 12} d^{\mu(12/d)}
=2^{\mu(6)}\,4^{\mu(3)}\,6^{\mu(2)}\,12^{\mu(1)}
=2\cdot 4^{-1}\cdot 6^{-1}\cdot 12
=1,\]
while $\operatorname{rad}(12)=2\cdot 3=6$.

\item \textbf{True.}
Let $\zeta_q=e^{2\pi i/q}$. The $q$th roots of unity split into disjoint sets of primitive $d$th roots as $d$ ranges over divisors of $q$.
Therefore
\[
\sum_{d\mid q} c_d(n)
=\sum_{\zeta^q=1} \zeta^n
=\begin{cases}q,& q\mid n,\\ 0,& q\nmid n,\end{cases}
\]
since it is the full geometric sum over $q$th roots.

\item \textbf{True.}
From the factorization $x^n-1=\prod_{d\mid n}\Phi_d(x)$ and Möbius inversion on the divisor lattice,
\[
\Phi_n(x)=\prod_{d\mid n}(x^d-1)^{\mu(n/d)}.
\]
\end{enumerate}

Thus $(t_1,t_2,t_3,t_4,t_5)=(1,1,0,1,1)$ and the Scroll Output is
\[
\boxed{11011}.
\]

\FinalAnswer{11011}

\NextPage
