% =========================
% puzzles/pXX.tex
% Final - The Gallery Grid of Half-True Plaques
% =========================

\PuzzleChapterIntro{The Gallery Grid of Half-True Plaques}{Logic Grid \textbullet\ Deduction \textbullet\ Unique Matching}{A logic-grid gallery with half-true plaques. Spatial constraints, time ordering, and five XOR inscriptions force a unique configuration.}



\Lore{

Five niches, five wards, five keepers.

Each alcove bears a two-line plaque: one sentence in brass, one in stone.
The Vault was built by skeptics, so every plaque lies once and tells truth once.

Tonight, the gallery is not merely locked.
A seized launch-capsule sits behind the eastern wall, and its countermand pad accepts \emph{five digits}---no more---before the failsafe seals.

The ledger that should have named the relics is scraped clean.
Only the half-true plaques remain, and the corridor will not forgive a single contradiction.

\textit{“Trust the grid. Then trust the order.”}

}



\Problem

\begingroup\scriptsize

Five niches lie in a straight row from west to east, labeled I, II, III, IV, V.
Each niche holds exactly one artifact, placed by exactly one custodian, at exactly one stamped time, sealed by exactly one ward.
Each artifact, custodian, time, and ward is used exactly once.

\medskip

\noindent\textbf{Artifacts:}
Obsidian Spindle; Ivory Lark; Jade Censer; Copper Astrolabe; Ashglass Urn.

\noindent\textbf{Custodians:}
Aegis; Athena; Kestrel; Quill; Sable.

\noindent\textbf{Times:}
21:11; 21:14; 21:17; 21:20; 21:23.

\noindent\textbf{Wards:}
Ash; Brass; Mirror; Salt; Thread.

\medskip

\noindent\textbf{Archivist’s fixed notes (all true).}
\begin{enumerate}[leftmargin=*]
\item The Mirror-warded niche is immediately to the west of the Ash-warded niche, and the Ash ward is on an \emph{end} niche (I or V).
\item The Thread ward is not on an end niche.
\item The Mirror ward was set at the latest time.
\item The Brass ward was set exactly three minutes after the Ash ward.
\item The Thread ward was set at a time \emph{strictly} between the Brass ward’s time and the Mirror ward’s time (so Brass $<$ Thread $<$ Mirror).
\item The Obsidian Spindle is the artifact sealed with Salt.
\item The Obsidian Spindle sits somewhere west of the Jade Censer.
\item The Jade Censer is \emph{not} in niche IV, and it is in a niche adjacent to the Copper Astrolabe.
\item The Ashglass Urn sits somewhere west of the Ivory Lark, and the Ivory Lark was placed earlier than the Jade Censer.
\item Exactly one of these two statements is true: \emph{Sable carried the Copper Astrolabe}; \emph{Kestrel carried the Ivory Lark}.
\item If Quill did \emph{not} carry the Obsidian Spindle, then Quill’s time was 21:14.
\item Sable did not use the Mirror ward and did not use the Thread ward.
\item In the signet registry, the names sort alphabetically as
Aegis \(\prec\) Athena \(\prec\) Kestrel \(\prec\) Quill \(\prec\) Sable.
The custodian who set the Mirror ward comes \emph{later} in that order than the custodian who set the Thread ward.
\end{enumerate}

\medskip

\noindent\textbf{The five plaques (each plaque has exactly one true line).}
\begin{enumerate}[leftmargin=*]
\item \textbf{Plaque One:}
\begin{itemize}
\item The Mirror ward stands immediately to the right of the Brass ward.
\item Aegis worked the middle niche (III).
\end{itemize}

\item \textbf{Plaque Two:}
\begin{itemize}
\item Athena set the Salt ward.
\item The Ash ward was set at 21:20.
\end{itemize}

\item \textbf{Plaque Three:}
\begin{itemize}
\item The Jade Censer is not in an end niche (I or V).
\item Quill carried the Ashglass Urn.
\end{itemize}

\item \textbf{Plaque Four:}
\begin{itemize}
\item The Thread ward is in niche II.
\item The Copper Astrolabe is in niche V.
\end{itemize}

\item \textbf{Plaque Five:}
\begin{itemize}
\item The Ivory Lark’s time was exactly six minutes earlier than the Thread ward’s time.
\item Aegis carried the Ivory Lark.
\end{itemize}
\end{enumerate}

\medskip

\noindent\textbf{Countermand protocol (how to extract the 5-digit PIN).}

After you determine the unique full assignment, order the \emph{five placements by time} from earliest to latest.
Let $r(\cdot)$ be the alphabetical rank from Note 13:
\[
r(\text{Aegis})=1,\ r(\text{Athena})=2,\ r(\text{Kestrel})=3,\ r(\text{Quill})=4,\ r(\text{Sable})=5.
\]
For each placement at time $21{:}mm$ by custodian $C$, form one digit as the \emph{units digit} of $mm+r(C)$.
Concatenate the five digits (earliest to latest) to obtain the PIN.

\endgroup

\VaultDirective{Determine the five-digit countermand PIN. (You will need to deduce the complete assignment of artifact, custodian, time, and ward to each niche.)}



\SolutionPage

\scriptsize

From Note 1, Ash is an end niche and Mirror is immediately to its west, so
\[
V=\text{Ash},\qquad IV=\text{Mirror}.
\]
By Note 3, \(t(IV)=21{:}23\). Notes 4--5 give \(t(\text{Brass})=t(\text{Ash})+3\) and \(t(\text{Brass})<t(\text{Thread})<21{:}23\), so \(t(\text{Ash})\in\{21{:}11,21{:}14\}\).

Plaque 2 is half-true. If \(t(\text{Ash})=21{:}20\) then Brass would be \(21{:}23\), leaving no time strictly between Brass and Mirror, contradicting Note 5; hence the second line of Plaque 2 is false and the first is true:
$
\textbf{Athena set Salt.}
$
By Note 6, Salt seals the Obsidian Spindle, so Athena carried the Spindle.

By Note 11, Quill did not carry the Spindle, hence \(t(\text{Quill})=21{:}14\). If \(t(\text{Ash})=21{:}11\), then \(t(\text{Brass})=21{:}14\) and Quill would be the Brass custodian; under Notes 7--9 this forces the Jade Censer into niche \(IV\), contradicting Note 8 (\(\text{Jade}\neq IV\)). Therefore
\[
t(\text{Ash})=21{:}14,\quad t(\text{Brass})=21{:}17,\quad t(\text{Thread})=21{:}20,\quad t(\text{Salt})=21{:}11,
\]
and in particular Quill is the Ash custodian in niche \(V\).

Plaque 4 is half-true. If the Copper Astrolabe were in \(V\), then by Note 8 the Jade Censer would be adjacent to \(V\) and hence in \(IV\), again contradicting Note 8. Thus the second line of Plaque 4 is false and the first is true:
\[
II=\text{Thread}.
\]

Plaque 1 is half-true and \(IV=\text{Mirror}\). Hence ``Mirror immediately right of Brass'' forces Brass in \(III\), so the other line is false:
\[
III=\text{Brass},\qquad \text{Aegis}\neq III.
\]

By Note 12, Sable used neither Mirror nor Thread, so Sable is not in \(IV\) or \(II\). With \(I\) and \(V\) occupied by Athena and Quill, this forces
\[
\text{Sable in }III.
\]

Plaque 5 is half-true. Since \(t(\text{Thread})=21{:}20\), its first line asserts \(t(\text{Ivory})=21{:}14\), i.e.\ Ivory lies in niche \(V\). If instead the second line were true (Aegis carried Ivory), then Ivory would not be in \(V\), contradicting the time forced by Note 9 together with the ward-time skeleton already fixed. Hence the second line is false and the first is true:
\[
V=\text{Ivory Lark}.
\]
So Quill carried the Ivory Lark, and in Note 10 the clause ``Kestrel carried the Ivory Lark'' is false; therefore the other clause is true:
\[
\text{Sable carried the Copper Astrolabe}.
\]
Thus niche \(III\) holds the Copper Astrolabe.

By Note 8, the Jade Censer is adjacent to Copper (in \(III\)) and not in \(IV\), so
\[
II=\text{Jade Censer},\qquad IV=\text{Ashglass Urn}.
\]

The remaining custodians for \(II\) and \(IV\) are Aegis and Kestrel. By Note 13, the Mirror custodian (niche \(IV\)) is later alphabetically than the Thread custodian (niche \(II\)), so
\[
II=\text{Aegis},\qquad IV=\text{Kestrel}.
\]

Therefore the unique assignment is
\[
\renewcommand{\arraystretch}{1.05}
\begin{array}{|c|l|l|c|l|}
\hline
\textbf{Niche} & \textbf{Artifact} & \textbf{Custodian} & \textbf{Time} & \textbf{Ward}\\
\hline
\text{I}   & \text{Obsidian Spindle} & \text{Athena}  & 21{:}11 & \text{Salt}\\
\text{II}  & \text{Jade Censer}      & \text{Aegis}   & 21{:}20 & \text{Thread}\\
\text{III} & \text{Copper Astrolabe} & \text{Sable}   & 21{:}17 & \text{Brass}\\
\text{IV}  & \text{Ashglass Urn}     & \text{Kestrel} & 21{:}23 & \text{Mirror}\\
\text{V}   & \text{Ivory Lark}       & \text{Quill}   & 21{:}14 & \text{Ash}\\
\hline
\end{array}
\]

\medskip

\noindent\textbf{PIN.}
Chronological order:
$21{:}11(\text{Athena}),\,21{:}14(\text{Quill}),\,21{:}17(\text{Sable}),\,21{:}20(\text{Aegis}),\,21{:}23(\text{Kestrel}).$

Units digit of minute$+$rank:
$11{+}2\to3,\ 14{+}4\to8,\ 17{+}5\to2,\ 20{+}1\to1,\ 23{+}3\to6,$
hence $\boxed{38216}$.

\FinalAnswer{38216}



\NextPage
