\PuzzleChapterIntro{The Jade Serpent}{Logic \textbullet\ Knights \& Knaves \textbullet\ Implication}{This is a classic ``liar logic'' puzzle wrapped in a whodunit. You must determine the unique truth-values of four suspects' statements to identify the single culprit. The key lies in finding a self-contradiction within a suspect's own testimony to break the symmetry.}
\Lore{

The Jade Serpent was lifted cleanly from its keep, leaving a perfect rectangle in the dust where its coils had rested.

Caldera’s courtyard is spare: hedge, stone, and an open stretch between the iron gate and the plinth at the center.
The gardeners keep that open ground lightly dusted with pale pumice so that even soft shoes leave a record.

Two records survive, both the kind the Vault trusts more than recollection.

Pinned under the display glass is the gardener’s slip:

\begin{quote}
\textit{22{:}25 — plinth brushed; idol present. \quad 22{:}35 — plinth brushed; idol absent.}
\end{quote}

And beside it, a lantern-plate from the courtyard camera (the flame-log that timestamps each exposure):

\begin{quote}
\textit{22{:}31 — exposure log. The serpent is visible on the plinth.}
\end{quote}

At 22{:}31 a clerk combed the pumice smooth.
At 22{:}35 the clerk found \emph{exactly one} continuous trail: from the gate threshold to the plinth and back again—and no other footprints anywhere after the 22{:}31 smoothing.

Above the gate hangs the hearing rule, cut into brass:

\begin{quote}
\textit{One tongue is inverted. Every sentence from that mouth is false. All others speak only truth.}
\end{quote}

}

\Problem

Exactly one of the four suspects is a \emph{universal liar} (every one of their statements below is false).
The other three suspects speak only truth.
Exactly one suspect stole the jade serpent.

From the Lore box, the serpent is on the plinth at 22{:}31 and gone by 22{:}35, and there is exactly one post-22{:}31 trail from the gate to the plinth and back.
For brevity, call this trail the \emph{plinth-run}. The thief is the person who made the plinth-run.

\medskip

\textbf{Suspects} \(\{\)\textbf{Athena}, \textbf{Quill}, \textbf{Kestrel}, \textbf{Sable}\(\}\) give these statements:

\medskip

\textbf{Athena}:

\begin{enumerate}[label=(\arabic*)]
\item \emph{``I left the courtyard at 22{:}30 and did not return.''}
\item Exactly one of these is true: \emph{``Quill made the plinth-run''} and \emph{``Kestrel made the plinth-run.''}
\item Sable is the universal liar \emph{if and only if} my 22{:}30 story is true.
\end{enumerate}

\medskip

\textbf{Quill}:

\begin{enumerate}[label=(\arabic*)]
\item If Athena’s 22{:}30 story is true, then Sable did not make the plinth-run.
\item If I made the plinth-run, then Athena’s 22{:}30 story is false.
\item Exactly one of these is true: \emph{``I made the plinth-run''} and \emph{``Sable is the universal liar.''}
\end{enumerate}

\medskip

\textbf{Kestrel}:

\begin{enumerate}[label=(\arabic*)]
\item Exactly one of these is true: \emph{``Quill made the plinth-run''} and \emph{``Athena’s 22{:}30 story is true.''}
\item Sable is the universal liar \emph{only if} I made the plinth-run.
\item Quill did not make the plinth-run.
\end{enumerate}

\medskip

\textbf{Sable}:

\begin{enumerate}[label=(\arabic*)]
\item Athena’s 22{:}30 story is false.
\item Exactly one of these is true: \emph{``I made the plinth-run''} and \emph{``Quill made the plinth-run.''}
\item Athena’s 22{:}30 story is false \emph{if and only if} Kestrel made the plinth-run.
\end{enumerate}

\medskip

\VaultDirective{Who stole the Jade Serpent?}

\SolutionPage

Let \(S\) be the proposition “Athena’s 22{:}30 story is true.”

From the Lore box, there is exactly one plinth-run after 22{:}31, and the thief is the person who made that plinth-run.

\medskip

\textbf{Sable cannot be a truth-teller.}

Assume Sable is \emph{not} the universal liar, so Sable’s statements are true.

Then Sable(1) implies \(S\) is false.
With \(S\) false, Sable(3) implies “Kestrel made the plinth-run” is true.

But if Kestrel made the (unique) plinth-run, then neither Sable nor Quill made it.
That makes Sable(2) false (it claims exactly one of “Sable made the plinth-run” and “Quill made the plinth-run” is true).

This contradicts the assumption that Sable’s statements are true.
Therefore Sable is the universal liar.

\medskip

\textbf{The thief is therefore: }

Since Sable is the universal liar, Sable(1) is false, so \(S\) is true.

Also Sable(3) is false. In Sable(3), the left side “\(S\) is false” is false (since \(S\) is true).
For a statement of the form “\(A\) if and only if \(B\)” to be false, \(A\) and \(B\) must differ.
Thus “Kestrel made the plinth-run” must be true.

Since the plinth-run identifies the thief, Kestrel stole the jade serpent.

\FinalAnswer{Kestrel}

\NextPage
