\PuzzleChapterIntro{The Staircase Registry}{Adversarial Search \textbullet\ Symmetrization \textbullet\ Recurrence}{A three-stage ledger puzzle: you must extract hidden digits from an adversarial “distance” oracle, then force an extremal friendship structure via symmetrization, and finally convert a staircase matching recurrence into a closed form. Expect an information bound, a structural extremal argument, and an exact enumeration that culminates in one large integer.}

\Lore{
A torn folio from the annex shows three headings, written in the same steady hand:

\textit{Echo. Feast. Bindings.}

Under \textit{Echo} are three probes and three even replies.
Under \textit{Feast} is a single line: \textit{“Make the weak quartets as many as possible.”}
Under \textit{Bindings} is the old staircase of wardens and seals.

In the margin: \textit{“If any step is guessed, the next one fails.”}
}

\Problem
\textbf{I. The Echo-Ruler.}
A witness secretly chooses an ordered triple \((u,v,w)\) with \(0\le u,v,w\le 9\).
A probe is a triple \((a,b,c)\) with \(0\le a,b,c\le 9\), and the reply is
\[
R(a,b,c)=|u+v-a-b|+|v+w-b-c|+|w+u-c-a|.
\]
Let \(M\) be the minimum number of probes needed to guarantee determining \((u,v,w)\) in the worst case.

The folio records that the auditor used \(M\) probes and obtained:
\[
R(0,0,0)=22,\qquad R(9,2,0)=14,\qquad R(4,0,7)=10.
\]

\medskip
\textbf{II. The Feast of Friends.}
Set \(N:=100+10u+v\).
Among \(N\) people, some pairs are friends. A \emph{weak quartet} is a set of four people containing
\emph{exactly one} friendship pair.
Let \(Q(N)\) be the maximum possible number of weak quartets over all friendship graphs on \(N\) vertices.
Let \(k\) be the number of connected components (cliques) in an extremal configuration attaining \(Q(N)\).

\medskip
\textbf{III. The Staircase Registry.}
There are \(15\) wardens \(W_1,\dots,W_{15}\) and \(29\) seals \(S_1,\dots,S_{29}\).
A \emph{binding} of size \(r\) is a set of \(r\) disjoint Warden--Seal pairs.
It is \emph{legal} if \(W_i\) may only bind to seals \(S_j\) with \(j\le 2i-1\).
Let \(A(15,r)\) be the number of legal bindings of size \(r\).
Define
\[
r:=M+k+w,\qquad V:=A(15,r)-Q(N).
\]

\medskip
\VaultDirective{Compute \(V\).}

\SolutionPage
\small

\textbf{I (Echo).}
Each reply is even and lies in \([0,54]\), so there are at most \(28\) possible replies; hence two probes give
\(\le 28^2=784<10^3\) outcomes, so \(M\ge 3\). The folio lists \(M\) probes and shows three, hence \(M=3\).

From \(R(0,0,0)=22\) we get \(u+v+w=11\).
For \((9,2,0)\),
\[
14=R(9,2,0)=|u{+}v{-}11|+|v{+}w{-}2|+|u{+}w{-}9|
= w+(9-u)+|u+w-9|,
\]
so \(14=2w\) if \(u+w\ge 9\), and \(14=18-2u\) if \(u+w<9\); thus \(w=7\) or \(u=2\).
For \((4,0,7)\),
\[
10=R(4,0,7)=|u{+}v{-}4|+|v{+}w{-}7|+|u{+}w{-}11|
=|7-w|+|4-u|+v.
\]
If \(w=7\) then \(u+v=4\) and \(R=|4-u|+v\in\{0,2,4,6,8\}\), contradiction; hence \(u=2\).
Then \(v+w=9\) and \(10=|7-w|+2+v=|v-2|+v+2\), giving \(v=5\), \(w=4\).
So \((u,v,w)=(2,5,4)\).

\medskip
\textbf{II (Feast).}
Now \(N=100+10u+v=125\).
By the copying/symmetrization inequality \(2q(G)\le q(G_x)+q(G_y)\), an extremal graph is a disjoint union of cliques.
With clique sizes \(a_1,\dots,a_n\) (\(\sum a_i=N\)),
\[
q(G)=\sum_{i=1}^n \binom{a_i}{2}\sum_{\substack{j<\ell\\ j,\ell\neq i}} a_j a_\ell.
\]
For fixed \(n\), \(q(G)\) is maximized when the nonzero \(a_i\) differ by \(\le 1\); for equal size \(a=N/n\),
\[
q_n=n\binom{a}{2}\binom{n-1}{2}a^2=\frac{N^3}{4}\cdot \frac{(n-1)(n-2)(N-n)}{n^3}.
\]
For \(N=125\), the factor \(f(n)=\frac{(n-1)(n-2)(125-n)}{n^3}\) satisfies
\(f(4)=\frac{726}{64}<f(5)=\frac{1440}{125}\) and \(f(6)=\frac{2380}{216}<f(5)\), so the maximum is at \(n=5\).
Since \(125=5\cdot 25\), an extremal configuration has \(k=5\) cliques of size \(25\), and
\[
Q(125)=5\binom{25}{2}\binom{4}{2}25^2=5\cdot 300\cdot 6\cdot 625=5{,}625{,}000.
\]

\medskip
\textbf{III (Registry).}
\(r=M+k+w=3+5+4=12\).
The staircase recurrence gives \(A(n,r)=(2n-r)A(n-1,r-1)+A(n-1,r)\), hence \(A(n,r)=\binom{n}{r}^2 r!\).
Thus
\[
A(15,12)=\binom{15}{3}^2\,12!=455^2\cdot 12!=99{,}165{,}306{,}240{,}000,
\]
so
\[
V=A(15,12)-Q(125)=99{,}165{,}306{,}240{,}000-5{,}625{,}000
=99{,}165{,}300{,}615{,}000.
\]

\normalsize\FinalAnswer{99165300615000}
\NextPage
